### 2022-06-22 docs/rmd/paper.tex
autogit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage[numbers]{natbib}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Analyzing the influence of Selection on Genetic Programming's
Generalization ability in Symbolic Regression:

A comparison of \(\epsilon\)-Lexicase Selection and Tournament Selection

--------------------------------------------------------

Student: Roman Höhn

Date of Birth: 1991-04-14

Place of Birth: Wiesbaden, Hesse

Student ID: 2712497

Supervisor: David Wittenberg

--------------------------------------------------------

Research Project for Seminar Information Systems (03.996.3299)

Chair of Business Administration and Computer Science

Johannes Gutenberg University Mainz

Summerterm 2022}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2022-06-22}

\begin{document}
\maketitle

\thispagestyle{empty}

\tableofcontents
\thispagestyle{empty}
\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Genetic programming (GP), a subfield of evolutionary computation (EC),
is a metaheuristic that is used to search for computer programs that
solve a given problem by simulating the process of darwinian evolution.
The basic principle of GP is to gradually evolve solutions by repeatedly
selecting parent solutions from a randomized population of computer
programs based on a fitness metric. Then, genetic operators are applied
on the selected solutions to genereate new offspring candidate
solutions. By repeating this process over many generations, GP acts as a
guided search for high fitness solutions throughout the decision space.
A unique feature of GP among other evolutionary optimization procedures
is the possibility to evolve solutions of variable length.

The overall performance of GP can depend strongly on the choice of its
underlying operators, one crucial component in this is the operator for
parent selection. Tournament Selection is a commonly used selection
operator in EC and is the most used operator in GP systems (Fang and Li,
2010, p. 181). A parent solution is selected by randomly sampling \(k\)
individuals from the current population into a tournament pool and then
the solution with the highest fitness score from the tournament pool is
selected (Fang and Li, 2010, p. 182). Lexicase selection has been
suggested as an alternative to tournament selection that is not based on
aggregating fitness scores. It samples \(n\) test cases in random order
and then eliminates solutions from the selection pool on a per test case
basis if they are not performing on an elite level (Helmuth, Spector and
Matheson, 2015, p. 1).

Since regular Lexicase selection has been shown to perform suboptimal on
continous-valued optimization problems, a modified variation called
\(\epsilon\)-Lexicase selection has been suggested for symbolic
regression by La Cava et. al (2016). Here, \(\epsilon\)-Lexicase
selection has shown itself to outperform both in overall performance
while showing only negligible computational overhead (La Cava, Spector
and Danai, 2016, p. 747).

The goal of symbolic regression is to find a mathematical model for an
observed set of datapoints (Paris, Robilliard and Fonlupt, 2004, p.
794). Symbolic Regression has been one of the first GP applications and
to this day is an actively studied and highly relevant area of research
(Poli, Langdon and McPhee, 2008, p. 114). In most symbolic regression
problems, little to no a priori knowledge about the optimal form and
structure of the target function is available. The ability of GP to
optimize for model structure as well as for parameters has lead to it
being one of the most prevalent methods used in the domain of symbolic
regression (Paris, Robilliard and Fonlupt, 2004, p. 795).

An important quality of all supervised machine learning applications,
including GP, is the ability to not only optimize performance for the
test cases a model is trained on but to also perform well on previously
unseen cases, this is refered to as generalization. In most real world
applications of symbolic regression only a small subset of labeled data
is available for training. The aim is to produce a model that not only
accurately predicts the provided training data but can also predict
previously unseen cases with high precision (Gonçalves, 2016, p. 6). A
model that is extensivley optimized on the provided training data, may
overfit to this data sample which may lead to a decrease in
generalization.

This research project tries to answer the question if the usage of
\(\epsilon\)-Lexicase selection influences the generalization behaviour
of programs that are evolved using GP for symbolic regression if
compared to programs that are evolved using traditional tournament
selection.

\hypertarget{an-overview-of-the-current-state-of-research}{%
\section{An Overview of the current state of
research}\label{an-overview-of-the-current-state-of-research}}

\hypertarget{selection}{%
\subsection{Selection}\label{selection}}

In the first description of lexicase\footnote{The alternative, more
  descriptive name given by Spector (2012): global pool, uniform random
  sequence, elitist lexicase parent selection} selection for GP by
Spector (2012) it was suggested as a novel parent selection method for
target problems that are modal in nature. The author classified modal
problems as ``problems that qualitatively different modes of response
are required for inputs from different regions of the problem's
domain''(Spector, 2012, p. 1).

In a following article Lexicase selection has been proposed specifically
for the purpose of solving so called uncompromising problems with GP.
Helmuth, Spector and Matheson (2015) defined uncompromising problems as
problems that require the final solution to perform optimal on each of
the cases it is tested on, examples include symbolic regression, the
design of digital multipliers or finding terms in finite algebras. The
authors also provided evidence that lexicase selection can significantly
improves GP's ability to solve some uncompromising problems if compared
to selection methods that are based on aggregated fitness (Helmuth,
Spector and Matheson, 2015, p. 12).

Both articles argued that many of the problem domains that GP is
commonly used for are prevalent for problems that are modal/
uncompromising in nature. In comparison to other selection operators,
populations that are evolved using variations of the lexicase selection
operator show a very high degree of genetic diversity which might be a
key contributer to the improved performance (Helmuth, Spector and
Matheson, 2015, p. 1) (La Cava, Spector and Danai, 2016, p. 745). In
theory, Lexicase based selection should be more prone to select
solutions that are specialist (high fitness on a small subset of
training cases) than solutions that are generalist (high average fitness
among all training cases).

The first description\footnote{Spector (2012), p.4} of the lexicase
parent selection algorithm by Spector (2012) is quoted below for
reference:

\begin{verbatim}
Lexicase - Parent-Selection:

1. Initialize:
  (a) Set candidates to be the entire population.
  (b) Set cases to be a list of all of the fitness cases in random order.

2. Loop:
  (a) Set candidates to be the subset of the current candidates that have 
      exactly the best fitness of any individual currently in candidates for 
      the first case in cases.
  (b) If candidates or cases contains just a single element then return
      the first individual in candidates.
  (c) Otherwise remove the first case from cases and go to Loop.
\end{verbatim}

More recent research by La Cava, Spector and Danai (2016) suggested
\(\epsilon\)-Lexicase selection as a modified selection operator that
can improve overall GP performance if applied to continous-valued
symbolic regression tasks in comparison to tournament and standard
lexicase selection. La Cava, Spector and Danai (2016) argued that the
original concept behind lexicase selection as formulated by Spector
(2012) does not fit the requirements of real-world symbolic regression
tasks. The authors identified the pass condition of regular lexicase
selection as the main problem if applied to symbolic regression:
Individual solutions can only be selected if they perform on an elite
level but with noisey and continous-valued data it is very unlikely for
two individuals to achieve an exactly equal error on a test case, this
finally results in filtering out too many individuals during the
selection process(La Cava, Spector and Danai, 2016, p. 742).
\(\epsilon\)-Lexicase adresses this problem by introducing an
\(\epsilon\) parameter that specifies a range around the elite error,
individuals that perform inside this range pass the current selection
iteration. La Cava \emph{et al.} (2017) describe \(\epsilon\)-Lexicase
selection as a more ``relaxed version of lexicase selection''. Different
methods to configure the \(\epsilon\) Parameter have been explored, for
the limited scope of this project I will focus on the most promising
implementation of \(\epsilon\)-Lexicase selection that is automatically
adjusted on the basis of the median absolute deviation of errors inside
the selection pool (La Cava, Spector and Danai, 2016, p. 742) (La Cava
\emph{et al.}, 2017, p. 6).

The performance increase of \(\epsilon\)-Lexicase selection for symbolic
regression problems has been demonstrated and reported for many
benchmark problems which led to widespread adaption of it in symbolic
regression applications (La Cava, Spector and Danai, 2016, pp.
744--745).

In comparison, traditional GP selection methods such as tournament
selection most commonly compute the fitness of a program as the mean of
its error for each individual fitness case. One downside to this
approach is that the total amount of information is reduced from a wide
range of individual errors to a single metric. Helmuth, Spector and
Matheson (2015) suspected that this loss in overall information provided
to GP might reduce overall performance especially if applied to the
class of uncompromising problems that require the solution to perform
well on a wide range of diverse cases. Since tournament selection
selects individuals based on their average fitness across all test
cases, the resulting solutions should be expected to be more biased
towards being generalists instead of specialists.

The basic algorithm for tournament is given below for comparison (Fang
and Li, 2010, pp. 182--183):

\begin{verbatim}
Tournament - Parent-Selection:

k = tournament size

1. Sample:
  (a) Randomly sample k individuals from the current population
      into a tournament pool.
2. Select:
  (a) Compute the mean fitness for each individual inside the tournament pool
      based on all fitness cases.
  (b) Select individual from the tournament pool with the highest mean fitness.
\end{verbatim}

\hypertarget{generalization}{%
\subsection{Generalization}\label{generalization}}

Generalization, the ability of a model to perform well on previously
unseen cases, is one of the fundamental goals in most real world machine
learning applications. O'Neill \emph{et al.} (2010) raised awareness to
the fact, that the topic of generalization in GP has not gotten the
attention that other machine learning-based areas, e.g, deep learning,
have attributed to it. A similiar statement was proposed by Kushchu
(2002) in his review of research in generalization in GP. Both authors
adressed the need for additional research regarding the topic.

Kushchu (2002) critized that almost all applications published in the
initial GP literature by John Koza (Koza, 1992) are not using separate
training and testing datasets which might result in overfitting and a
poor overall generalization ability of the programs that are produced.
The author calls for a more widespread adoption of methods like
generational sampling of new training cases or the overall separation
into training and testing datasets to improve generalization in GP
(Kushchu, 2002, p. 10).

An aspect that is suspected to negatively correlate with GP's
generalization is overall size and in particular bloating of the
resulting programs. Bloating can be described as a growth in the total
size of a program that does not also improve it's performance in any
meaningful way (Silva and Vanneschi, 2009, p. 1). It has been widely
suspected that GP configurations that produce very large programs also
have a higher tendency to specialize on difficult or unusual test cases
which could result in lower generalization (Wang, Wagner and Rondinelli,
2019, p. 268). The minimum description length principle (MDLP) describes
this phenomenom by claiming that less complex models are more likely to
perform better at generalizing than more complex models that achieve a
comparable fitness during the training phase (O'Neill \emph{et al.},
2010, p. 349). However MDLP should be considered carefully, experiments
by Silva and Vanneschi (2009) on the prediction of bioavailability for
drugs demonstrated that GP based techniques can either bloat without
overfitting or overfit without bloating (Silva and Vanneschi, 2009, p.
8).

\hypertarget{symbolic-regression}{%
\subsection{Symbolic Regression}\label{symbolic-regression}}

The task of finding a mathematical function that fits a given set of
datapoints has been one of the first applications of GP in (Koza, 1992,
p. 238). Practical examples given by Koza (1992) included tasks such as
the discovery of scientific laws, finding solutions to differential and
integral equations or the programmatic compression of images.

Since it's inception, GP based symbolic regression has been widely used
in practical application ranging over a wide field of disciplines. A
well known example of an industrial use-case has been published by
Jordaan \emph{et al.} (2004). The authors used GP based symbolic
regression to derive nonlinear functions that could be used to
significantly increase the robustness of industrial sensors which
resulted in wide-spread adoption and a significant reduction in costs.

Many variations of GP based symbolic regression have been proposed to
improve overall performance, Wang, Wagner and Rondinelli (2019)
summarized the main characteristics and differences for 6 common
variations including procedures such as geometric semantic genetic
programming, cartesian genetic programming or GP-based relevance vector
machines. A large benchmarking study on symbolic regression published by
Orzechowski, Cava and Moore (2018) compared 4 GP-based procedures to
several state-of-the-art machine leaning techniques. The authors find
that GP, eventhough more time consuming, can achieve better results if
compared to the state-of-the-art machine learning algorithm
gradient-boosting.

\hypertarget{experimental-study}{%
\section{Experimental study}\label{experimental-study}}

\hypertarget{research-design}{%
\subsection{Research Design}\label{research-design}}

To study the influence of selection on generalization I selected a
dataset about energy efficiency in buildings that is part of the UC
Irvine Machine Learning Repository (Dua and Graff, 2017). The dataset
contains eight individual building attributes that map to two different
outcomes, heating and cooling load, for N=768 cases (Tsanas and Xifara,
2012).

Using two otherwise identical GP systems, one deploying tournament
selection and the other \(\epsilon\)-lexicase selection, the objective
is to find a computer program that best predicts the outcome variable
heating load \((Y1)\)\footnote{The symbolic regression is performed on
  one of the two provided outcome variables, the variable Cooling Load
  will be excluded.} of the buildings using a subset of the eight
building attributes \((X1,..,X8)\) for input. The specific meaning of
all attributes inside the dataset are described in table 1.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-1}Overview - Energy Heating Dataset}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Variable} & \textbf{Description}\\
\hline
X1 & Relative Compactness\\
\hline
X2 & Surface Area\\
\hline
X3 & Wall Area\\
\hline
X4 & Roof Area\\
\hline
X5 & Overall Height\\
\hline
X6 & Orientation\\
\hline
X7 & Glazing Area\\
\hline
X8 & Glazing Area Distribution\\
\hline
y1 & Heating Load\\
\hline
y2 & Cooling Load\\
\hline
\end{tabular}
\end{table}

To measure the generalization ability of each model the dataset will be
randomly split in half, resulting in a training and testing dataset each
containing 384 individual cases. Each model will be evolved by
traditional GP using only the fitness cases present in the training
dataset. For each generation the highest fitness model of the current
population will be tested with the previously unseen fitness cases that
are part of the testing dataset. For each run of the experiment
statistics will be collected on fitness which will form the basis of my
further statistical analysis. Additional statistics will be collected
for the average length of the population and the length of the elite
program for each generation to explore differences in their distribution
and possible correlations to generalization.

Since GP is a stochastic optimization algorithm, the basic experiment
will be run for a total of 50 times to ensure a fair and meaningful
comparison based on a large number of runs for both algorithms. For each
run of the experiment the dataset will be randomly split in half as
described, both models are then trained and tested using the exact same
set of fitness cases.

The statistical analysis of the collected data will first focus on
examining the question if, on average, the usage of
\(\epsilon\)-Lexicase selection will result in models that perform
significantly different than models that are evolved using tournament
selection both.

\begin{quote}
\(H0_{1}\): The distribution underlying the samples of training/testing
errors produced by torunament selection is the same as the distribution
underlying samples of training/testing errors produced by
\(\epsilon\)-lexicase selection
\end{quote}

The next question of interest is, to examine if statistical significant
differences between the mean errors of training and testing data exist
for both selection operators:

\begin{quote}
\(H0_{2}\): The distribution underlying the samples of training errors
produced by \(\epsilon\)-Lexicase/torunament selection is the same as
the distribution underlying samples of testing errors produced by
\(\epsilon\)-Lexicase/tournament selection
\end{quote}

To gather further insight into the differences between both GP systems,
an additional test will be performed on the question if differences in
the total size of the resulting programs exist:

\begin{quote}
\(H0_{3}\): Size differences exist between the distribution underlying
the samples produced by \(\epsilon\)-Lexicase and the distribution
underlying samples of tournament selection
\end{quote}

All hypothesis will be tested using a level of significance of
\(\alpha=0.05\).

To further examine the difference in generalization behaviour, the mean
testing and training errors over each generation will be visualized for
both algorithms. The specific aim of this visualization is to explore
the mechanism of overfitting and to examine if differences between both
methods can be detected.

All GP experiments will be implemented by using the python programming
language in conjunction with DEAP, a framework for distributed
evolutionary algorithms that implements various tools and algorithms for
genetic programming (Fortin \emph{et al.}, 2012).

\hypertarget{gentetic-programming-configuration}{%
\subsection{Gentetic Programming
Configuration}\label{gentetic-programming-configuration}}

\hypertarget{evolutionary-parameters}{%
\subsubsection{Evolutionary parameters}\label{evolutionary-parameters}}

The basic evolutionary parameters for both systems are presented in
table 2. Tournament selection is used with a default tournament size of
\(3\) individuals while the \(\epsilon\) parameter in lexicase selection
is selected automatically as previously mentioned in subsection 3.1.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-2}Evolutionary Parameters}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Parameter} & \textbf{Value}\\
\hline
Population Size & 500\\
\hline
Number of Generations & 100\\
\hline
Mutation Rate & 20\%\\
\hline
Crossover Rate & 80\%\\
\hline
Tournament Size & 3\\
\hline
Epsilon selection & automatic\\
\hline
Elite Size & 0\\
\hline
\end{tabular}
\end{table}

\hypertarget{fitness-evaluation}{%
\subsubsection{Fitness Evaluation}\label{fitness-evaluation}}

The fitness \(f\) for each model will be based on the mean squared error
(MSE) over all fitness cases for prediction and the empirically measured
values as described by eq.1.

\begin{equation}
\tag{eq. 1}
MSE = \frac{1}{n} * \sum_{i=1}^{n} (Y_i - \hat{Y_1})^2
\end{equation}

where:

\begin{itemize}
\tightlist
\item
  \(n\): total number of test cases
\item
  \(Y_i\): empirical value for case \(i\)
\item
  \(\hat{Y_1}\): predicted value for case \(i\).
\end{itemize}

The resulting fitness function \(f\) for an individual program \(i\) is
descibed by eq 2:

\begin{equation}
\tag{eq. 2}
 f(i, \tau ) = \frac{1}{N} * \sum_{t \epsilon \tau} (y_t - y\hat{}_t(i, x_t))^2 
\end{equation}

where\footnote{Naming and notation was adopted from La Cava, Spector and
  Danai (2016)}:

\begin{itemize}
\tightlist
\item
  \(\tau\): the set of \(N\) fitness cases
\item
  \(y_t\): empirical value of the target for case \(t\)
\item
  \(y\hat{}_t(i, x_t)\): predicted value for the target for case \(t\)
  by running the program \(i\) with the total set of input variables
  \(x_t\)
\end{itemize}

\hypertarget{primitive-set}{%
\subsubsection{Primitive Set}\label{primitive-set}}

The primitve set consists of the terminals that are listed in table 3
and the functions that are listed in table 4. To avoid runtime errors I
implemented protected version of the operators for division, natural
logarithm and square root (Koza, 1992, pp. 82--83). As suggested by Koza
(1992) I also included ephemeral constants to the set of terminals to
provide the evolutionay search the opportunity to explore and include
randomly generated constants.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Terminals}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Terminal} & \textbf{Description}\\
\hline
X1 & Relative Compactness\\
\hline
X2 & Surface Area\\
\hline
X3 & Wall Area\\
\hline
X4 & Roof Area\\
\hline
X5 & Overall Height\\
\hline
X6 & Orientation\\
\hline
X7 & Glazing Area\\
\hline
X8 & Glazing Area Distribution\\
\hline
random\_int & Ephemeral Constant (integer)\\
\hline
random\_float & Ephemeral Constant(float)\\
\hline
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-4}Functions}
\centering
\begin{tabular}[t]{l|r}
\hline
\textbf{Function} & \textbf{Arity}\\
\hline
Addition & 2\\
\hline
Subtraction & 2\\
\hline
Multiplication & 2\\
\hline
Negation & 1\\
\hline
Sine & 1\\
\hline
Cosine & 1\\
\hline
Protected Division & 2\\
\hline
Protected Natural Logarithm & 1\\
\hline
Protected Square Root & 1\\
\hline
\end{tabular}
\end{table}

\hypertarget{genetic-operators}{%
\subsubsection{Genetic Operators}\label{genetic-operators}}

GP operators are represented in table 5. Both genetic operators use a
static limit to control for the height of the reulting trees (Koza,
1992, p. 104). Individual programs are initialized by using the ramped
half-and-half method, 50\% of the population are created by using the
Growth algorithm and the remaining 50\% are created by using the Full
algorithm (Koza, 1992, p. 93).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-5}GP Operators}
\centering
\begin{tabular}[t]{l|l|r}
\hline
\textbf{Operator} & \textbf{Implementation} & \textbf{Static.Height.Limit}\\
\hline
Initilization & Ramped Half/Half & 2\\
\hline
Crossover & One Point Crossover & 17\\
\hline
Mutation & Uniform Mutation & 17\\
\hline
\end{tabular}
\end{table}

The crossover operator implemented by DEAP randomly selects a crossover
point in each individual and exchanges each subtree with the point as
root between each individual (Fortin \emph{et al.}, 2012). Mutation also
randomly selects a point in the tree individual, it then replaces the
subtree below that point as a root by the expression generated using the
full grow initialization method (Fortin \emph{et al.}, 2012).

\hypertarget{results}{%
\section{Results}\label{results}}

Tables 6 and 7 summarize the results for all fitness scores collected
over 50 total runs of the experiment\footnote{All floating point numbers
  in tables have been rounded to 3 decimal places if not stated
  otherwise}.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-6}Summary - Tournament Selection}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrrrr}
\toprule
\textbf{X} & \textbf{gen} & \textbf{nevals} & \textbf{mean\_training\_error} & \textbf{std\_training\_error} & \textbf{min\_training\_error} & \textbf{max\_training\_error} & \textbf{testing\_error} & \textbf{std\_testing\_error} & \textbf{avg\_size} & \textbf{elite\_size}\\
\midrule
count & 101.0 & 101.000 & 1.010000e+02 & 1.010000e+02 & 101.000 & 1.010000e+02 & 100.000 & 100.000 & 100.000 & 100.000\\
mean & 50.0 & 420.742 & 2.092614e+18 & 3.305357e+20 & 16.483 & 1.046309e+21 & 16.905 & 28.858 & 55.674 & 62.070\\
std & 29.3 & 8.115 & 2.098028e+19 & 3.313954e+21 & 12.514 & 1.049016e+22 & 11.117 & 15.732 & 34.504 & 36.850\\
min & 0.0 & 417.300 & 2.582030e+08 & 1.463390e+10 & 7.066 & 1.193720e+11 & 7.847 & 16.217 & 3.331 & 4.340\\
25\% & 25.0 & 418.780 & 2.151894e+10 & 3.044239e+12 & 8.479 & 1.070232e+13 & 9.350 & 18.344 & 25.133 & 28.195\\
\addlinespace
50\% & 50.0 & 420.000 & 3.494635e+11 & 5.511534e+13 & 11.918 & 1.747285e+14 & 12.874 & 22.715 & 57.299 & 66.400\\
75\% & 75.0 & 421.000 & 4.892936e+12 & 6.455802e+14 & 18.773 & 2.238973e+15 & 19.685 & 32.750 & 87.852 & 96.625\\
max & 100.0 & 500.000 & 2.108540e+20 & 3.330558e+22 & 77.243 & 1.054272e+23 & 62.693 & 95.282 & 109.761 & 115.340\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-7}Summary - Epsilon-Lexicase Selection}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrrrr}
\toprule
\textbf{X} & \textbf{gen} & \textbf{nevals} & \textbf{mean\_training\_error} & \textbf{std\_training\_error} & \textbf{min\_training\_error} & \textbf{max\_training\_error} & \textbf{testing\_error} & \textbf{std\_testing\_error} & \textbf{avg\_size} & \textbf{elite\_size}\\
\midrule
count & 101.0 & 101.000 & 1.010000e+02 & 1.010000e+02 & 101.000 & 1.010000e+02 & 100.000 & 100.000 & 100.000 & 100.000\\
mean & 50.0 & 420.781 & 7.533781e+12 & 1.186094e+15 & 22.013 & 3.762638e+15 & 22.456 & 39.048 & 58.416 & 63.819\\
std & 29.3 & 8.109 & 5.900311e+13 & 9.313627e+15 & 16.670 & 2.950203e+16 & 15.904 & 19.225 & 34.515 & 36.356\\
min & 0.0 & 416.240 & 4.660413e+05 & 3.314323e+07 & 9.680 & 1.539125e+08 & 11.129 & 21.487 & 3.219 & 4.280\\
25\% & 25.0 & 418.900 & 2.528193e+08 & 2.727451e+10 & 11.316 & 1.192373e+11 & 12.089 & 24.860 & 28.233 & 29.615\\
\addlinespace
50\% & 50.0 & 420.040 & 2.196352e+09 & 2.647194e+11 & 14.478 & 1.098068e+12 & 15.223 & 30.057 & 64.689 & 71.440\\
75\% & 75.0 & 421.020 & 2.027020e+11 & 3.198174e+13 & 23.511 & 1.012976e+14 & 23.305 & 51.081 & 87.394 & 94.910\\
max & 100.0 & 500.000 & 5.788911e+14 & 9.137151e+16 & 74.622 & 2.894465e+17 & 77.518 & 110.244 & 108.785 & 115.660\\
\bottomrule
\end{tabular}}
\end{table}

Some interesting observation from tables 6 and 7 can be made by
comparing the mean of the variables \(min\_training\_error\) and
\(testing\_error\). Variable \(min\_training\_error\) represents the
fitness, measured by the MSE, of the best performing solution that has
been found during a single run of the experiment while the variable
\(testing\_error\) represents the fitness of the same solution if
evaluated for the unknown fitness cases inside the testing dataset.
Tournament selection based GP achieves a mean \(min\_training\_error\)
of \(16.483\) and a mean \(testing\_error\) of \(16.905\) while the GP
system using \(\epsilon\)-Lexicase selection can only achieve a mean
\(min\_training\_error\) of \(22.013\) and a mean \(testing\_error\) of
\(22.456\). These results indicate an overall performance benefit for
torunament selection based GP in the experiment. It should also be
mentioned that the best performing model over all 50 runs, measured by
the minimum of variables \(min\_training\_error\) and
\(testing\_error\), has been achieved by tournament selection with an
MSE of 7.066 for the training data and 7.847 for the testing data.

If we compare the differences of \(min\_training\_error\) and
\(testing\_error\) on the basis of each algorithm, we can observe that
both tournament and \(\epsilon\)-lexicase achieve a slightly lower error
on the training data than on the testing data. If we compute the total
difference between the means of \(min\_training\_error\) and
\(testing\_error\) for both algorithms tournament selection amounts to
0.422 and \(\epsilon\)-Lexicase selection to 0.443. Regarding our
primary research question, the differences in generalization, the
initial observations do not seem to indicate a significant difference in
the the relative gap between the solutions performance on testing data
and training data.

Figure 1 visualizes the distribution of \(min\_training\_error\) and
\(testing\_error\) for both selection operators over the total 50 runs
using a boxplot and summerizes the initial observations from tables 6
and 7 described above.

\begin{figure}
\centering
\includegraphics{./plots/mean_error_boxplot_all.png}
\caption{Distribution of Errors}
\end{figure}

The four samples from figure 1 have been tested for normal distribution
by computing the D'Agostino and Pearson test (D'Agostino, 1971)
(D'Agostino and Pearson, 1973) to determine if they satisfy the
requirements of the students t-test. The results are summarized in table
8. Since not all samples test positive for a normal distribution a
Mann-Whitney U (MWU) ranksum test will be performed to test for
statistical significance between the four variables
\(tournament\_min\_training\_error\), \(tournament\_testing\_error\),
\(e-lexicase\_min\_training\_error\) and \(e-lexicase\_testing\_error\).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-8}Normal Distribution Tests}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrl}
\toprule
\textbf{sample} & \textbf{statistic} & \textbf{p.value} & \textbf{alpha} & \textbf{normal\_distributed}\\
\midrule
tournament\_min\_training\_error & 36.146 & 0.000 & 0.05 & False\\
e-lexicase\_min\_testing\_error & 1.575 & 0.455 & 0.05 & True\\
tournament\_testing\_error & 32.503 & 0.000 & 0.05 & False\\
e-lexicase\_testing\_error & 59.135 & 0.000 & 0.05 & False\\
\bottomrule
\end{tabular}}
\end{table}

The results of the MWU test are summarized in table 9.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-9}Mean Error - P-Values}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrr}
\toprule
\textbf{ } & \textbf{tournament\_training\_errors} & \textbf{tournament\_testing\_errors} & \textbf{elexicase\_training\_errors} & \textbf{elexicase\_testing\_errors}\\
\midrule
tournament\_training\_errors & 1.000 & 0.309 & 0.000 & 0.000\\
tournament\_testing\_errors & 0.309 & 1.000 & 0.002 & 0.000\\
elexicase\_training\_errors & 0.000 & 0.002 & 1.000 & 0.257\\
elexicase\_testing\_errors & 0.000 & 0.000 & 0.257 & 1.000\\
\bottomrule
\end{tabular}}
\end{table}

The mean overall performance of tournament selection appears to be
superior to that of \(\epsilon\)-Lexicase selection which could be
considered contradictory evidence to the \ldots{} interpretation\ldots{}

To further examine the generalization behaviour based on both selection
operators I plotted the mean fitness scores for both algorithms in
figure 2.

\begin{figure}
\centering
\includegraphics{./plots/mean_error_combined.png}
\caption{Mean Errors}
\end{figure}

\ldots{} interpretation\ldots{}

In a final step I analyzed the growth behaviour for both GP systems.
Figure 3 visualizes the average size of individuals inside the
population as well as the average size of the best performing individual
for each generation of the evolution.

\includegraphics{./plots/size_combined.png} \ldots{}
interpretation\ldots{}

Again, a MWU test is conducted to test for statistical differences in
the underlying distribution of the sizes measured during the experiment.
The results are summarized in table 10.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-10}Mean Size - P-Values}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrr}
\toprule
\textbf{ } & \textbf{tournament\_elite\_size} & \textbf{elexicase\_elite\_size} & \textbf{tournament\_avg\_size} & \textbf{elexicase\_avg\_size}\\
\midrule
tournament\_elite\_size & 1.000 & 0.858 & 0.533 & 0.652\\
elexicase\_elite\_size & 0.858 & 1.000 & 0.764 & 0.682\\
tournament\_avg\_size & 0.533 & 0.764 & 1.000 & 0.992\\
elexicase\_avg\_size & 0.652 & 0.682 & 0.992 & 1.000\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{limitations-and-open-questions}{%
\section{Limitations and open
questions}\label{limitations-and-open-questions}}

I faced two major difficulties in the preparation and evaluation of this
research project:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Configuration of evolutionary parameters
\item
  Limited computational ressources
\end{enumerate}

The combination of both issues might have resulted in a significant
reduction in overall robustness of the results represented in sections 5
and 6. The performance of all evolutionary algorithms, including GP, can
be highly dependant on a large number of parameters and implementation
details. Although the series of experiments conducted in this project
are comparably simple in nature, the results achieved might still be
highly influenced by the GP configuration detailed in subsection 4.2.

To gain further insight into the differences in generalizaion behaviour
and to adress the research question with a higher level of certainty,
the experiment I conducted should be repeated for many different GP
configurations. Some exemplary parameters that could drastically
influence the results of this experiment include the population size,
the total number of generations, the inclusion of elitism or the
computation of \(\epsilon\) in lexicase selection. Other factors whose
influence on generalization behaviour could be important include
different methods to compute an individuals fitness (e.g.~mean absolute
error or root mean squared error) and testing out different variations
of the genetic operators for mutation and crossover.

Another obvious weakness of my project is that the experiment is only
based on a single symbolic regression application, the prediction of the
heating load of buildings based on a relativley small dataset. To gain
further trust in the obtained results, the experiment should be repeated
for other symbolic regression tasks and datasets. At the current state
it remains unknown if the generalization behaviour of both selection
operators might be highly problem specific and dependant on the dataset
used for training and testing.

All limitations and shortcomings described above are also connected to
the second limitation of this research project, the limited amount of
computational ressources. The total computation time consumed to run the
current experiments (50 runs, 2 algorithms, 100 generations, 500
individuals) has already been close to 24 hours on a modern, arm-based
Apple-M1 system. Eventhough the overall computation time could probably
be reduced by further optimization of the source code, e.g.~using tools
for parallel computation, the ressources necessary to repeat the whole
experiment for different parameter configurations would certainly be
beyond the scope of this research project.

\newpage

\hypertarget{I}{%
\section*{I References}\label{I}}
\addcontentsline{toc}{section}{I References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-DAgostino1971AnOT}{}}%
D'Agostino, R.B. (1971) {``An omnibus test of normality for moderate and
large size samples,''} \emph{Biometrika}, 58, pp. 341--348.

\leavevmode\vadjust pre{\hypertarget{ref-DAgostino1973TestsFD}{}}%
D'Agostino, R.B. and Pearson, E.S. (1973) {``Tests for departure from
normality,''} in.

\leavevmode\vadjust pre{\hypertarget{ref-Dua:2019}{}}%
Dua, D. and Graff, C. (2017) {``{UCI} machine learning repository.''}
University of California, Irvine, School of Information; Computer
Sciences. Available at: \url{http://archive.ics.uci.edu/ml}.

\leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-642-16493-4_19}{}}%
Fang, Y. and Li, J. (2010) {``A review of tournament selection in
genetic programming,''} in Cai, Z. et al. (eds.) \emph{Advances in
computation and intelligence}. Berlin, Heidelberg: Springer Berlin
Heidelberg, pp. 181--192.

\leavevmode\vadjust pre{\hypertarget{ref-DEAP_JMLR2012}{}}%
Fortin, F.-A. \emph{et al.} (2012) {``{DEAP}: Evolutionary algorithms
made easy,''} \emph{Journal of Machine Learning Research}, 13, pp.
2171--2175.

\leavevmode\vadjust pre{\hypertarget{ref-Gonalves2016AnEO}{}}%
Gonçalves, I. (2016) {``An exploration of generalization and overfitting
in genetic programming: Standard and geometric semantic approaches,''}
in.

\leavevmode\vadjust pre{\hypertarget{ref-6920034}{}}%
Helmuth, T., Spector, L. and Matheson, J. (2015) {``Solving
uncompromising problems with lexicase selection,''} \emph{IEEE
Transactions on Evolutionary Computation}, 19(5), pp. 630--643.
doi:\href{https://doi.org/10.1109/TEVC.2014.2362729}{10.1109/TEVC.2014.2362729}.

\leavevmode\vadjust pre{\hypertarget{ref-soft_sensor_gp}{}}%
Jordaan, E. \emph{et al.} (2004) {``Robust inferential sensors based on
ensemble of predictors generated by genetic programming,''} in Yao, X.
et al. (eds.) \emph{Parallel problem solving from nature - PPSN VIII}.
Berlin, Heidelberg: Springer Berlin Heidelberg, pp. 522--531.

\leavevmode\vadjust pre{\hypertarget{ref-koza_main}{}}%
Koza, J.R. (1992) \emph{Genetic programming: On the programming of
computers by means of natural selection}. Cambridge, MA, USA: MIT Press.
Available at: \url{http://mitpress.mit.edu/books/genetic-programming}.

\leavevmode\vadjust pre{\hypertarget{ref-generalisation_in_gp}{}}%
Kushchu, I. (2002) {``An evaluation of EvolutionaryGeneralisation in
genetic programming,''} \emph{Artificial Intelligence Review - AIR}, 18,
pp. 3--14.
doi:\href{https://doi.org/10.1023/A:1016379201230}{10.1023/A:1016379201230}.

\leavevmode\vadjust pre{\hypertarget{ref-https:ux2fux2fdoi.orgux2f10.48550ux2farxiv.1709.05394}{}}%
La Cava, W. \emph{et al.} (2017) {``A probabilistic and multi-objective
analysis of lexicase selection and epsilon-lexicase selection.''} arXiv.
doi:\href{https://doi.org/10.48550/ARXIV.1709.05394}{10.48550/ARXIV.1709.05394}.

\leavevmode\vadjust pre{\hypertarget{ref-epsilon_lexicase_main}{}}%
La Cava, W., Spector, L. and Danai, K. (2016) {``Epsilon-lexicase
selection for regression,''} in \emph{Proceedings of the genetic and
evolutionary computation conference 2016}. New York, NY, USA:
Association for Computing Machinery (GECCO '16), pp. 741--748.
doi:\href{https://doi.org/10.1145/2908812.2908898}{10.1145/2908812.2908898}.

\leavevmode\vadjust pre{\hypertarget{ref-open_issues_gp}{}}%
O'Neill, M. \emph{et al.} (2010) {``Open issues in genetic
programming,''} \emph{Genetic Programming and Evolvable Machines}, 11,
pp. 339--363.
doi:\href{https://doi.org/10.1007/s10710-010-9113-2}{10.1007/s10710-010-9113-2}.

\leavevmode\vadjust pre{\hypertarget{ref-Orzechowski_2018}{}}%
Orzechowski, P., Cava, W.L. and Moore, J.H. (2018) {``Where are we
now?''} in \emph{Proceedings of the genetic and evolutionary computation
conference}. {ACM}.
doi:\href{https://doi.org/10.1145/3205455.3205539}{10.1145/3205455.3205539}.

\leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
Paris, G., Robilliard, D. and Fonlupt, C. (2004) {``Exploring
overfitting in genetic programming,''} in Liardet, P. et al. (eds.)
\emph{Artificial evolution}. Berlin, Heidelberg: Springer Berlin
Heidelberg, pp. 267--277.

\leavevmode\vadjust pre{\hypertarget{ref-poli08:fieldguide}{}}%
Poli, R., Langdon, W.B. and McPhee, N.F. (2008) \emph{A field guide to
genetic programming}. Published via \texttt{http://lulu.com}; freely
available at \texttt{http://www.gp-field-guide.org.uk}. Available at:
\url{https://digitalcommons.morris.umn.edu/cgi/viewcontent.cgi?article=1001\&context=cs_facpubs}.

\leavevmode\vadjust pre{\hypertarget{ref-bloat_overfitting_gp}{}}%
Silva, S. and Vanneschi, L. (2009) {``Operator equalisation, bloat and
overfitting: A study on human oral bioavailability prediction,''} in
\emph{Proceedings of the 11th Annual Genetic and Evolutionary
Computation Conference, GECCO-2009}, pp. 1115--1122.
doi:\href{https://doi.org/10.1145/1569901.1570051}{10.1145/1569901.1570051}.

\leavevmode\vadjust pre{\hypertarget{ref-lexicase_first_desciption}{}}%
Spector, L. (2012) {``Assessment of problem modality by differential
performance of lexicase selection in genetic programming: A preliminary
report.''}
doi:\href{https://doi.org/10.1145/2330784.2330846}{10.1145/2330784.2330846}.

\leavevmode\vadjust pre{\hypertarget{ref-fe8fa39e88a040bbacba5a465c48043f}{}}%
Tsanas, A. and Xifara, A. (2012) {``Accurate quantitative estimation of
energy performance of residential buildings using statistical machine
learning tools,''} \emph{Energy and buildings}, 49, pp. 560--567.
doi:\href{https://doi.org/10.1016/j.enbuild.2012.03.003}{10.1016/j.enbuild.2012.03.003}.

\leavevmode\vadjust pre{\hypertarget{ref-wang_wagner_rondinelli_2019}{}}%
Wang, Y., Wagner, N. and Rondinelli, J.M. (2019) {``Symbolic regression
in materials science,''} \emph{MRS Communications}, 9(3), pp. 793--805.
doi:\href{https://doi.org/10.1557/mrc.2019.85}{10.1557/mrc.2019.85}.

\end{CSLReferences}

\newpage

\hypertarget{II}{%
\section*{II Statutory Declaration}\label{II}}
\addcontentsline{toc}{section}{II Statutory Declaration}

\end{document}