### 2022-06-23 docs/rmd/paper.Rmd
autogit

---
title: |
  <center>Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression: </center>
  <center>
    A comparison of $\epsilon$-lexicase Selection and Tournament Selection
  </center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Research Project for Seminar Information Systems (03.996.3299)</center>
  <center>FB 03: Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  <center>Summerterm 2022</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    includes:
      in_header: settings.tex
# indent: true
fontsize: 12pt
# toc: yes
# toc_depth: 4
bibliography: bib/bib/ref.bib
csl: bib/csl/harvard1.csl
# include-before: '`\newpage{} `{=latex}'
include-before: '`\thispagestyle{empty} \newpage{}`{=latex}'
---

```{r setup, include=FALSE}
library(kableExtra)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H")
```

\tableofcontents
\thispagestyle{empty}
\newpage

\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

# Introduction

```{=html}
<!-- 
Structure:
  - Genetic Programming
  - Selection
  - Symbolic Regression
  - Generalization
  - Research Question
-->
```

<!-- GP --->
Genetic programming (GP), a sub field of evolutionary computation (EC), is a meta heuristic that is used to search for computer programs that solve a given problem by simulating the process of Darwinian evolution. The basic principle of GP is to gradually evolve solutions by repeatedly selecting parent solutions from a randomized population of computer programs based on a fitness metric. Then, genetic operators are applied on the selected solutions to generate new offspring candidate solutions. By repeating this process over many generations, GP acts as a guided search for high fitness solutions throughout the decision space.
A unique feature of GP among other evolutionary optimization procedures is the possibility to evolve solutions of variable length.

<!-- Selection --->
The overall performance of GP can depend strongly on the choice of its underlying operators, one crucial component in this is the operator for parent selection. 
Tournament Selection is a commonly used selection operator in EC and is the most used operator in GP systems [@10.1007/978-3-642-16493-4_19, p.181]. A parent solution is selected by randomly sampling $k$ individuals from the current population into a tournament pool and then the solution with the highest fitness score from the pool is selected [@10.1007/978-3-642-16493-4_19, p.182]. 
Lexicase selection has been suggested as an alternative to tournament selection that is not based on aggregating fitness scores. It samples $n$ test cases in random order and then eliminates solutions from the selection pool on a per test case basis if they are not performing on an elite level [@6920034, p.1].

Since regular Lexicase selection has been shown to perform sub optimal on continuous-valued optimization problems, a modified variation called $\epsilon$-lexicase selection has been suggested for symbolic regression by La Cava et. al (2016). Here, $\epsilon$-lexicase selection has shown itself to outperform both in overall performance while showing only negligible computational overhead [@epsilon_lexicase_main, p.747].

<!-- Symbolic Regression --->
The goal of symbolic regression is to find a mathematical model for an observed set of data points [@10.1007/978-3-540-24621-3_22, p.794]. Symbolic Regression has been one of the first GP applications and to this day is an actively studied and highly relevant area of research [@poli08:fieldguide, p.114]. In most symbolic regression problems, little to no prior knowledge about the optimal form and structure of the target function is available. The ability of GP to optimize for model structure as well as for parameters has lead to it being one of the most prevalent methods used in the domain of symbolic regression [@10.1007/978-3-540-24621-3_22, p.795].

<!-- Generalization --->
An important quality of all supervised machine learning applications, including GP, is the ability to not only optimize performance for the test cases a model is trained on but to also perform well on previously unseen cases, this is referred to as generalization. In most real world applications of symbolic regression only a small subset of labeled data is available for training. The aim is to produce a model that not only accurately predicts the provided training data but can also predict previously unseen cases with high precision [@Gonalves2016AnEO, p.6]. A model that is extensively optimized on training data, can become over fitted to the data sample which may lead to a decrease in generalization.

<!-- Research Question --->
This research project tries to answer the question if the usage of $\epsilon$-lexicase selection influences the generalization behavior of programs that are evolved using GP for symbolic regression if compared to programs that are evolved using traditional tournament selection.

<!-- Vorgehen --->
Section 2 will give a brief summary and discussion of topic related scientific research. Section 3 describes the design of the experiment as well as GP specific implementation details. The statistical  evaluation of all relevant results and the summarized conclusions are given in section 3 and 4. Section 5 addresses limitation and open questions of this research project.

# An Overview of the current state of research

## Selection

In the first description of lexicase^[The alternative, more descriptive name given by @lexicase_first_desciption: global pool, uniform random sequence, elitist lexicase parent selection] selection for GP by @lexicase_first_desciption it was suggested as a novel parent selection method for target problems that are modal in nature. The author classifies modal problems as "problems that qualitatively different modes of response are required for inputs from different regions of the problem's domain"[@lexicase_first_desciption, p.1].

In a following article Lexicase selection has been proposed specifically for the purpose of solving so called uncompromising problems with GP. @6920034 defines uncompromising problems as problems that require the final solution to perform optimal on each of the cases it is tested on, examples include symbolic regression, the design of digital multipliers or finding terms in finite algebras. The authors also provide evidence that lexicase selection can significantly improves GP's ability to solve some uncompromising problems if compared to selection methods that are based on aggregated fitness [@6920034, p.12].

Both articles argue that many of the problem domains that GP is commonly used for are prevalent for problems that are modal or uncompromising in nature. In comparison to other selection operators, populations that are evolved using variations of the lexicase selection operator show a very high degree of genetic diversity which might be a key contributor to the improved performance [@6920034, p.1] [@epsilon_lexicase_main, p.745]. In theory, Lexicase based selection should be more prone to select solutions that are specialist (solutions that achieve high fitness on a smaller subset of unusual training cases) than solutions that are generalist (solutions that achieve high average fitness among all training cases).

The first description of the lexicase parent selection algorithm is quoted verbatim below for reference [@lexicase_first_desciption, p.4]:

```
Lexicase - Parent-Selection:

1. Initialize:
  (a) Set candidates to be the entire population.
  (b) Set cases to be a list of all of the fitness cases in random order.

2. Loop:
  (a) Set candidates to be the subset of the current candidates that have 
      exactly the best fitness of any individual currently in candidates for 
      the first case in cases.
  (b) If candidates or cases contains just a single element then return
      the first individual in candidates.
  (c) Otherwise remove the first case from cases and go to Loop.
```

More recent research by @epsilon_lexicase_main suggests $\epsilon$-lexicase selection as a modified selection operator that can improve overall GP performance if applied to continuous-valued symbolic regression tasks in comparison to tournament and standard lexicase selection.  @epsilon_lexicase_main argues that the original concept behind lexicase selection as formulated by @lexicase_first_desciption does not fit the requirements of real-world symbolic regression tasks. The authors identifies the pass condition of regular lexicase selection as the main problem if applied to symbolic regression: Individual solutions can only be selected if they perform on an elite level but with continuous-valued, and often noisy, data it is very unlikely for two individuals to achieve an exact equal error for a given test case. This stringent pass condition can result in filtering out too many individuals during the selection process [@epsilon_lexicase_main, p.742]. $\epsilon$-lexicase adresses this problem by introducing an $\epsilon$ parameter that specifies a range around the elite error, individuals that perform inside this range pass the current selection iteration. @https://doi.org/10.48550/arxiv.1709.05394 describes $\epsilon$-lexicase selection as a more "relaxed version of lexicase selection".
Different methods to configure the $\epsilon$ Parameter have been explored, for the limited scope of this project I will focus on the most promising implementation of $\epsilon$-lexicase selection that is automatically adjusted on the basis of the median absolute deviation of errors inside the selection pool [@epsilon_lexicase_main, p.742] [@https://doi.org/10.48550/arxiv.1709.05394, p.6].

The performance increase of $\epsilon$-lexicase selection for symbolic regression problems has been demonstrated and reported for many benchmark problems which led to widespread adaption of it in symbolic regression applications [@epsilon_lexicase_main, p.744-745].

In comparison, a traditional GP selection method like tournament selection most commonly computes the fitness of a program as the mean of its error for each individual fitness case. One downside to this approach is that the total amount of information is reduced from a wide range of individual errors to a single metric. @6920034 suspects that this loss in overall information provided to GP might reduce overall performance especially if applied to the class of uncompromising problems that require the solution to perform well on a wide range of diverse cases. Since tournament selection selects individuals based on their average fitness across all test cases, the resulting solutions should be expected to be more biased towards being generalists instead of specialists.

The basic algorithm for tournament is given below for comparison [@10.1007/978-3-642-16493-4_19, p.182-183]:

```
Tournament - Parent-Selection:

k = tournament size

1. Sample:
  (a) Randomly sample k individuals from the current population
      into a tournament pool.
2. Select:
  (a) Compute the mean fitness for each individual inside the tournament pool
      based on all fitness cases.
  (b) Select individual from the tournament pool with the highest mean fitness score.
```

## Generalization

Generalization, the ability of a model to perform well on previously unseen cases, is one of the fundamental goals in most real world machine learning applications. @open_issues_gp raises awareness to the fact, that the topic of generalization in GP has not gotten the attention that other machine learning-based areas attributed to it. A similar statement has been proposed by @generalisation_in_gp in his review of research in generalization in GP. Both authors address the need for additional research regarding the topic.

@generalisation_in_gp criticizes that almost all applications published in the initial GP literature by @koza_main are not using separate training and testing data sets which might result in overfitting and a poor overall generalization ability of the programs that are produced. The author calls for a more widespread adoption of methods like generational sampling of new training cases or the overall separation into training and testing data sets to improve generalization in GP [@generalisation_in_gp, p.10].

An aspect that is suspected to negatively correlate with GP's generalization is overall size and in particular bloating of the resulting programs. Bloating can be described as a growth in the total size of a program that does not improve its performance in any meaningful way [@bloat_overfitting_gp, p. 1].
It has been widely suspected that GP configurations that produce very large programs also have a higher tendency to specialize on difficult or unusual test cases which could result in lower generalization [@wang_wagner_rondinelli_2019, p. 268]. The minimum description length principle (MDLP) describes this phenomenon by claiming that less complex models are more likely to perform better at generalizing than more complex models that achieve a comparable fitness during the training phase [@open_issues_gp, p. 349]. However MDLP should be considered with caution, experiments by @bloat_overfitting_gp on the prediction of bio availability for drugs demonstrate that GP based techniques can either bloat without overfitting or over fit without bloating [@bloat_overfitting_gp, p. 8].

## Symbolic Regression

The task of finding a mathematical function that fits a given set of data points has been one of the first applications of GP in [@koza_main, p. 238].
Practical examples given by @koza_main included tasks such as the discovery of scientific laws, finding solutions to differential and integral equations or the programmatic compression of images.

Since its inception, GP based symbolic regression has been widely used in practical applications ranging over a wide field of disciplines. A well known example of an industrial use-case has been published by @soft_sensor_gp. The authors use GP based symbolic regression to derive nonlinear functions that can be used to significantly increase the robustness of industrial sensors. This resulted in wide-spread adoption and a significant reduction in costs.

Many variations of GP based symbolic regression have been proposed to improve overall performance, @wang_wagner_rondinelli_2019 summarizes the main characteristics and differences for 6 common variations including procedures such as geometric semantic genetic programming, Cartesian genetic programming or GP-based relevance vector machines. A large benchmarking study on symbolic regression published by @Orzechowski_2018 compares 4 GP-based procedures to several state-of-the-art machine leaning techniques. The authors demonstrate that GP, even though more time consuming, can achieve better results if compared to the state-of-the-art machine learning algorithm gradient-boosting.

# Experimental study

## Research Design

<!-- data set and Task --->
To study the influence of selection on generalization I selected a data set about energy efficiency in buildings that is part of the UC Irvine Machine Learning Repository [@Dua:2019]. The data set contains eight individual building attributes that map to two different outcomes, heating and cooling load, for N=768 cases [@fe8fa39e88a040bbacba5a465c48043f].

Using two otherwise identical GP systems, one deploying tournament selection and the other $\epsilon$-lexicase selection, the objective is to find a computer program that best predicts the outcome variable heating load $(Y1)$^[The symbolic regression is performed on one of the two provided outcome variables, the variable Cooling Load will be excluded.] of the buildings using a subset of the eight building attributes $(X1,..,X8)$ for its input. The specific meaning of all attributes inside the data set are described in table 1.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"

knitr::kable(
  read.csv("./tables/csv/variables_energy_heating.csv"),
  digits=3,
  caption="Overview - Energy Heating data set"
)    %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

<!-- Method:single run --->
To measure the generalization ability of each model the data set will be randomly split in half, resulting in a training and testing data set each containing 384 individual cases. Each model will be evolved by traditional GP using only the fitness cases present in the training data set. For each generation the highest fitness model of the current population will be tested with the previously unseen fitness cases that are part of the testing data set. For each run of the experiment statistics will be collected on fitness which will form the basis of my further statistical analysis. Additional statistics will be collected for the average length of the population and the length of the elite program for each generation to explore differences in their distribution and possible correlations to generalization.

<!-- Statistical Analysis --->
Since GP is a stochastic optimization algorithm, the basic experiment will be run for a total of 50 times to ensure a fair and meaningful comparison based on a large number of runs for both algorithms. For each run of the experiment the data set will be randomly split in half as described, both models are then trained and tested using the exact same set of fitness cases.

The statistical analysis of the collected data will first focus on examining the question if, on average, the usage of $\epsilon$-lexicase selection will result in models that perform significantly different than models that are evolved using tournament selection:

> $H0_{1}$: The distribution underlying the samples of training/testing errors produced by torunament selection is the same as the distribution underlying samples of training/testing errors produced by $\epsilon$-lexicase selection

The next question of interest is, to examine if statistical significant differences between the mean errors of training and testing data exist for both selection operators:

> $H0_{2}$: The distribution underlying the samples of training errors produced by $\epsilon$-lexicase/torunament selection is the same as the distribution underlying samples of testing errors produced by $\epsilon$-lexicase/tournament selection

To gather further insight into the differences between both GP systems, an additional test will be performed on the question if differences in the total size of the resulting programs exist:

> $H0_{3}$: No differences in program size exist between the distribution underlying the samples produced by $\epsilon$-lexicase and the distribution underlying samples of tournament selection

All hypothesis will be tested using a level of significance of $\alpha=0.05$.

To further examine the difference in generalization behavior, the mean testing and training errors over each generation will be visualized for both algorithms. The specific aim of this visualization is to explore the mechanism of overfitting and to examine if differences between both methods can be detected.

All GP experiments will be implemented by using the python programming language in conjunction with DEAP, a framework for distributed evolutionary algorithms that implements various tools and algorithms for genetic programming [@DEAP_JMLR2012]. A repository that contains the full source code, tables, plots and other materials represented in this thesis is available at github.com^[https://github.com/roman91DE/intelligent_information_systems_research_project].

## Genetic Programming Configuration

### Evolutionary parameters

The basic evolutionary parameters for both systems are presented in table 2. Tournament selection is used with a default tournament size of $3$ individuals while the $\epsilon$ parameter in lexicase selection is selected automatically as previously mentioned in subsection 3.1.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"

knitr::kable(
  read.csv("./tables/csv/gp_parameters.csv"),
  digits=3,
  caption="Evolutionary Parameters"
) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

### Fitness Evaluation

The fitness $f$ for each model will be based on the mean squared error (MSE) over all fitness cases for prediction and the empirically measured values as described by eq.1.

\begin{equation}
\tag{eq. 1}
MSE = \frac{1}{n} * \sum_{i=1}^{n} (Y_i - \hat{Y_1})^2
\end{equation}

where:

-   $n$: total number of test cases
-   $Y_i$: empirical value for case $i$
-   $\hat{Y_1}$: predicted value for case $i$.

The resulting fitness function $f$ for an individual program $i$ is described by eq 2:

\begin{equation}
\tag{eq. 2}
 f(i, \tau ) = \frac{1}{N} * \sum_{t \epsilon \tau} (y_t - y\hat{}_t(i, x_t))^2 
\end{equation}

where^[Naming and notation was adopted from @epsilon_lexicase_main]:

-   $\tau$: the set of $N$ fitness cases
-   $y_t$: empirical value of the target for case $t$
-   $y\hat{}_t(i, x_t)$: predicted value for the target for case $t$ by running the program $i$ with the total set of input variables $x_t$

### Primitive Set

The primitive set consists of the terminals listed in table 3 and the functions listed in table 4. To avoid run time errors I implemented protected version of the operators for division, natural logarithm and square root [@koza_main, p.82-83]. As suggested by @koza_main I also included ephemeral constants to the set of terminals to provide the evolutionary search the opportunity to explore and include randomly generated constants.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"

knitr::kable(
  read.csv("./tables/csv/terminals.csv"),
  digits=3,
  caption="Terminals"
) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"

knitr::kable(
  read.csv("./tables/csv/primitives.csv"),
  digits=3,
  caption="Functions"
) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

### Genetic Operators

GP operators are represented in table 5. Both genetic operators use a static limit to control for the height of the resulting trees [@koza_main, p.104]. Individual programs are initialized by using the ramped half-and-half method, 50% of the population are created by using the Growth algorithm and the remaining 50% are created by using the Full algorithm [@koza_main, p.93].

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"

knitr::kable(
  read.csv("./tables/csv/genetic_operators.csv"),
  digits=3,
  caption="GP Operators"
) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

The crossover operator implemented by DEAP randomly selects a crossover point in each individual and exchanges each sub tree with the point as root between each individual [@DEAP_JMLR2012]. Mutation also randomly selects a point in the tree individual, it then replaces the sub tree below that point as a root by the expression generated using the full grow initialization method [@DEAP_JMLR2012].

# Results

Tables 6 and 7 summarize the results for all fitness scores collected over 50 total runs of the experiment^[All floating point numbers in tables have been rounded to 3 decimal places if not stated otherwise].

```{r, echo=FALSE, warning=FALSE}

fig.pos="H"

df <- read.csv("./tables/csv/master_tournament_descriptive.csv")

knitr::kable(
  df,
  format = "latex",
  booktabs = TRUE,
  digits=3,
  caption="Summary - Tournament Selection"
) %>% kable_styling(latex_options = "scale_down") %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

```

```{r, echo=FALSE, warning=FALSE}

fig.pos="H"

df <- read.csv("./tables/csv/master_elexicase_descriptive.csv")

knitr::kable(
  df,
  format = "latex",
  booktabs = TRUE,
  digits=3,
  caption="Summary - Epsilon-Lexicase Selection"
)  %>% kable_styling(latex_options = "scale_down") %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

```

A few interesting observations from tables 6 and 7 can be made by comparing the mean of the variables $min\_training\_error$ and $testing\_error$. Variable $min\_training\_error$ represents the fitness, measured by the MSE, of the best performing solution that has been found during a single run of the experiment. The variable $testing\_error$ represents the fitness of the same solution if evaluated for the unknown fitness cases inside the testing data set. Tournament selection based GP achieves a mean $min\_training\_error$ of $16.483$ and a mean $testing\_error$ of $16.905$ while the GP system using $\epsilon$-lexicase selection can only achieve a mean $min\_training\_error$ of $22.013$ and a mean $testing\_error$ of $22.456$. These results indicate an overall performance benefit for the tournament selection based GP in the experiment. It should also be mentioned that the best performing model over all 50 runs, measured by the minimum of variables $min\_training\_error$ and $testing\_error$, has been achieved by tournament selection with an MSE of 7.066 for the training data and	7.847 for the testing data.

If we compare the differences of $min\_training\_error$ and $testing\_error$ on the basis of each algorithm, we can observe that both tournament and $\epsilon$-lexicase achieve a slightly lower fitness on the training data than on the testing data. We compute the total difference between the means of $min\_training\_error$ and $testing\_error$ for both algorithms tournament selection amounts to 0.422 and $\epsilon$-lexicase selection to 0.443. The initial observations do not seem to indicate a significant difference in the the relative gap between the solutions' performance on testing data and training data.

Figure 1 visualizes the distribution of $min\_training\_error$ and $testing\_error$ for both selection operators over the total 50 runs using a box plot and summarizes the initial observations from tables 6 and 7 described above.

![Distribution of Errors](./plots/mean_error_boxplot_all.png)

The four samples from figure 1 have been tested for normal distribution by computing the D’Agostino and Pearson test [@DAgostino1971AnOT] [@DAgostino1973TestsFD] to determine if they satisfy the requirements of the students t-test. The results are summarized in table 8. Since not all samples test positive for a normal distribution a Mann-Whitney U (MWU) rank sum test will be performed to test for statistical significance between the four variables $tournament\_min\_training\_error$, $tournament\_testing\_error$, $e-lexicase\_min\_training\_error$ and $e-lexicase\_testing\_error$.

```{r, echo=FALSE, warning=FALSE}

fig.pos="H"

df <- read.csv("./tables/csv/normal_dist_test.csv")

knitr::kable(
  df,
  format = "latex",
  booktabs = TRUE,
  digits=3,
  caption="Normal Distribution Tests"
) %>% kable_styling(latex_options = "scale_down") %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

```

Table 9 displays the resulting p-values from computing the MWU test in a matrix format.

```{r, echo=FALSE, warning=FALSE}

fig.pos="H"

df <- read.csv("./tables/csv/mwu_matrix_error.csv")

knitr::kable(
  df,
  format = "latex",
  booktabs = TRUE,
  digits=3,
  caption="Mean Error - P-Values"
) %>% kable_styling(latex_options = "scale_down") %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

```

The results show two major findings:

1. The differences in mean fitness between tournament selection and $\epsilon$-lexicase selection are highly statistical significant even if considering an $\alpha$ of 0.01. For the given symbolic regression task tournament selection achieves a better mean performance on both the training as well as the testing data set than $\epsilon$-lexicase selection. These findings contradict the hypothesis $H0_{1}$ described in subsection 3.1 and are contrary to some of the research that was summarized in subsections 2.1 and 2.3.[@epsilon_lexicase_main] [@https://doi.org/10.48550/arxiv.1709.05394].

2. The differences in mean fitness between $min\_training\_error$ and $testing\_error$ are not statistically significant for neither selection algorithms. When comparing the mean training error to the mean testing error, the p-values only compute to 0.309 for tournament selection and to 0.257 for $\epsilon$-lexicase selection. These findings fail to reject $H0_{2}$, based on the obtained results we can not find evidence that both algorithms demonstrate a significantly different performance between the training data and the previously unseen testing data.

To further examine the generalization behavior of both selection operators I plotted the mean fitness scores for both algorithms over the total number of 100 generations in figure 2. Figure 2 shows that both algorithms show little visual evidence of overfitting which would result in a growing gap between between $min\_training\_error$ and $testing\_error$ in the final stages of the evolutionary optimization. For $\epsilon$-lexicase selection a small gap can be observed starting approximately at the 90th generation, while the minimum training error slowly converges a small increase in the testing error. Based on the limited amount of collected data, it is questionable if this small sign of overfitting can be interpreted as a weak indicator for an inferior generalization behavior of $\epsilon$-lexicase in comparison to tournament selection.

Figure 2 also shows a visual representation of the described overall performance benefit of using tournament selection instead of $\epsilon$-lexicase selection for the chosen symbolic regression task. It can be observed that at all stages of the evolution tournament selection based GP achieves a lower mean error for both data sets.

![Mean Errors](./plots/mean_error_combined.png)

In a final step I analyze the growth behavior for both GP systems to determine if differences in program size can be observed based on selection operators. Figure 3 visualizes the average size of individuals inside the population as well as the average size of the best performing individual for each generation and both GP systems. 
\begin{figure}
\centering
\includegraphics{./plots/size_subplotted.png}
\caption{Mean Size}
\end{figure}

<!-- ![Mean Size](./plots/size_sub plotted.png)--->

Both subplots display a very similar behavior that is common for GP. We observe a strong trend for solutions to grow in size with each additional generation that is concluded. Another observation  that can be made is that the best solutions of each generation almost always have a larger size than the population average. On first impression, the differences between both selection operators appear to be insignificant. Again, I conducted a MWU test to examine statistical differences in the underlying distribution of sizes measured during the experiment.

The resulting p-values of the MWU test are summarized in table 10.

```{r, echo=FALSE, warning=FALSE}
fig.pos="H"

df <- read.csv("./tables/csv/mwu_matrix_size.csv")

knitr::kable(
  df,
  format = "latex",
  booktabs = TRUE,
  digits=3,
  caption="Mean Size - P-Values"
) %>% kable_styling(latex_options = "scale_down") %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

```

None of the differences in mean size are statistically significant. Again, based on the results obtained in my experiment, the evidence produced fails to reject the null hypothesis $H0_{3}$ that selection has no significant effect on differences in program size. It should be noted that the configuration of both GP algorithms did include two mechanisms that place limitations on the growth behavior of the individual solutions to avoid excessive bloating (see sub section 3.2.4).

# Conclusion

The objective of this research project is to answer the question if the choice between two commonly used parent selection operators, tournament selection and $\epsilon$-lexicase selection, significantly influences the generalization behavior of GP for symbolic regression tasks.

The results obtained from my experiment do not demonstrate any statistically significant differences in generalization. Both GP systems achieve a slightly lower mean error if evaluated for the training data sets than if evaluated for the unseen fitness cases of the testing data set. The mean differences between testing and training errors do not show statistically significant differences for neither algorithms (summarized in table 9).

To further explore differences in generalization behavior the mean for both algorithms training and testing error over all generations for all 50 independent runs is visualized and analyzed (see figure 2). A very small and probably insignificant gap does show for $\epsilon$-lexicase selection in the final phase of the evolutionary search. Comprehensively, this approach also does not yield any clear evidence of differences in generalization.

A final attempt to explore possible differences between both GP systems is to explore differences in growth behavior as an indicator for generalization. This idea is based on the wide spread hypothesis that increases in overall program size might be an indication of overfitting to the training data. This concept has been discussed in sub section 2.2. The results do not show statistically significant differences in the growth behavior between both selection operators.

An interesting finding of this analysis is, that the overall performance of tournament selection shows to be notably higher than the performance of $\epsilon$-lexicase selection for the selected symbolic regression problem. This finding is rather unexpected based on the research of @epsilon_lexicase_main and @https://doi.org/10.48550/arxiv.1709.05394 that is the primary source of inspiration for this research project.

# Limitations and open questions

I faced two major difficulties in the preparation and evaluation of this research project:

  1. Configuration of evolutionary parameters
  2. Limited computational resources

The combination of both issues might have resulted in a significant reduction in overall robustness of the results represented in sections 5 and 6. The performance of all evolutionary algorithms, including GP, can be highly dependent on a large number of parameters and implementation details. Although the series of experiments conducted in this project are comparably simple in nature, the results achieved might still be highly influenced by the GP configuration detailed in subsection 4.2.

To gain further insight into the differences in generalization behavior and to address the research question with a higher level of certainty, the experiment I conducted should be repeated for different GP configurations. Some exemplary parameters that could drastically influence the results of this experiment include the population size, the total number of generations, the inclusion of elitism or the computation of $\epsilon$ in lexicase selection. Other factors whose influence on generalization behavior could be important include different methods to compute an individuals fitness (e.g. mean absolute error or root mean squared error), testing different variations of the genetic operators for mutation and crossover and experimenting with different variations of the training:testing data set split size.

Another obvious weakness of this project is that the experiment is only based on a single symbolic regression application, the prediction of the heating load of buildings based on a relatively small data set. To gain further trust in the obtained results, the experiment should be repeated for other symbolic regression tasks and data sets. It remains unknown if the generalization behavior of both selection operators might be highly problem specific and dependent on the data set used for training and testing.

All limitations and shortcomings described above are also connected to the second limitation of this research project, the limited amount of computational resources. The total computation time consumed to run the current experiments (50 runs, 2 algorithms, 100 generations, 500 individuals) has already been close to 24 hours on a modern, arm-based Apple-M1 system. Even though the overall computation time could probably be reduced by further optimization of the source code, e.g. using tools for parallel computation, the resources necessary to repeat the whole experiment for different parameter configurations would certainly be beyond the scope of this research project.

\newpage

# I - References {.unnumbered #I}

::: {#refs}
:::

# II - Statutory Declaration {.unnumbered #II}

![Statutory Declaration](./private/erklaerung.pdf)\